---
title: "project 4"
author: "Qiang"
date: "2016年7月6日"
output: html_document
---

# quality of white wines exploration by Xia

* dataset Overview : 

    This  data set contains 4,898 white wines with 11 variables on quantifying the chemical properties of each wine. At least 3 wine experts rated the quality of each wine, providing a rating between 0 (very bad) and 10 (very excellent).

* I want to know:

    Which chemical properties influence the quality of white wines?
    
    May I predict quality rating through some chemical properties?
* Things I will do

    look data distribution in these 11 variables 
    
    make correlation matrix to pick up some varibale with relative larger correlation
    
    figure out how these variables influence quality 
    
    build a model that may predict quality by these variables
    
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```


##  1 data distribution

```{r}
# load data
library(ggplot2)
library(reshape2)
library(magrittr)
library(dplyr)
library(plyr)
wqw <- read.csv("wineQualityWhites.csv")
str(wqw)
names(wqw)
```

### quality

```{r}
### quality rating is discrete values(5, 6, 7...),we factor it before plot
ggplot(data = wqw,aes(as.factor(quality)))+ 
  geom_bar()
table(wqw$quality)
```

most wines quality was between 5 and 7 

### volatile.acidity
```{r}
## creat histogram function 
create_plot <- function(variable, binwidth = 0.01) {
  return(ggplot(aes_string(x = variable), data = wqw) + 
           geom_histogram(binwidth = binwidth))}
create_plot("volatile.acidity",binwidth=0.01)
summary(wqw$volatile.acidity)
```

distribution of volatile.acidity is slightly right skew
most wines' volatile acidity content are between 0.21 and 0.32, median 0.26 and mean 0.2782 . and there some outliers that have high content of  volatile acidity


### free.sulfur.dioxide
```{r}

create_plot("free.sulfur.dioxide",binwidth=3)
summary(wqw$free.sulfur.dioxide)
```


most wines' free.sulfur.dioxide content are between 20 and 50 , median 34 and mean 35.31 ,there are and there some outliers that have high content of free sulfur dioxide. so let's zoom in

```{r}
ggplot(wqw,aes(free.sulfur.dioxide))+
  geom_histogram(binwidth = 1)+
  coord_cartesian(xlim = c(0,quantile(wqw$free.sulfur.dioxide,0.99)))
```

the distribution of wines in free.sulfur.dioxide was mainly symmetry 

### citric.acid

```{r}

create_plot("citric.acid",binwidth=0.03)
summary(wqw$citric.acid)
```

the distribution of wine in citric.acid was almostly symmetry .
most wines have citric acid between 0.25 and 0.4, and there some higher outliers.




### residual.sugar
```{r}

create_plot("residual.sugar",binwidth=1)
summary(wqw$residual.sugar)
```
the distribution of wine in residual.sugar was right skew.
most wines' residual sugar are lower than 6,mean is 6,391, median is 5.2.
and there some outliers even higer than 20 .


### chlorides
```{r}
create_plot("chlorides",binwidth=0.001)
summary(wqw$chlorides)
```
most wines' chlorides contents are between 0.03 and 0.06 , mean 0.04577, median 0.043
and there are some higher content outlier 

let's zoom in
```{r}
ggplot(wqw,aes(chlorides)) + geom_histogram(binwidth = 0.003)+
  coord_cartesian(xlim = c(0,0.1))
```
the distribution of wine in chlorides was symmetry. 

### total.sulfur.dioxide
```{r}
create_plot("total.sulfur.dioxide",binwidth=10)
summary(wqw$total.sulfur.dioxide)
```
the distribution of wine in total.sulfur.dioxidechlorides was symmetry.
most wines' total sulfur.dioxide contents are between 100 and 180 , mean is 138.4, median  is 134

### density 
```{r}
create_plot("density",binwidth=0.001)
summary(wqw$density)

```
the distribution of wine in density was  symmetry , most density are between 0.98 and 1.00


### pH
```{r}
create_plot("pH",binwidth=0.1)
summary(wqw$pH)
```
the distribution of wine in Ph was symmetry. mean ph is 3.188 , median is 3.18.

### sulphates
```{r}
create_plot("sulphates",binwidth=0.05)
summary(wqw$sulphates)
```

the distribution of wine in sulphates was slightly right skewed. mean content is 0.4898, median is 0.47 .

### alcohol

```{r}
create_plot("alcohol",binwidth=0.1)
summary(wqw$alcohol)
```

the distribution of wine in alcohol was multimodal.
mean is 10.51 mg/L ,median is 11.40 mg/L







## 2 correlation matrix 

```{r }

## creat correlation matrix 
library(GGally)
ggcorr(wqw[, 2:13], hjust = 1, size = 3, color = "grey50",
       palette = "RdGy", label = TRUE, label_size = 3, label_color = "black")
```

* we find these variable have relatively higher correlation : 

1. alcohol
2. density
3. volatile.acidity
4. chlorides
5. total.sulfur.dioxide

* I want to know how these variables influence quality

## 3 Bivariate Analysis

#### alcohol Vs quanlity

look at alcohol distribute in wine with different quality

```{r}
ggplot(aes(factor(quality), 
           alcohol ),
       data = wqw) +
    geom_boxplot( alpha = .5)+
    geom_jitter( alpha = .3)  +
    geom_smooth(aes(quality-2, 
                    alcohol),
                method = "lm", 
                se = FALSE,size=2)
## One Way Anova 
 summary(aov(quality ~alcohol, data=wqw))

```
quality in 5~9, better wine tend to have higher content of alcohol.

One Way Anova show that , alcohol content do have significantly influence on quality rating.

#### density and quanlity


```{r}
ggplot(aes(factor(quality), 
            density), 
        data = wqw) +
    geom_boxplot( alpha = .5)+
    geom_jitter( alpha = .3)  +
    geom_smooth(aes(quality-2, 
                    density),
                method = "lm", 
                se = FALSE,size=2)
 summary(aov(quality ~density, data=wqw))
```

lower density winer tender to have higher quality.

One Way Anova show that : density have significantly influence on wine quanlity


#### volatile.acidity and quanlity

```{r}
ggplot(aes(factor(quality), 
            volatile.acidity), 
        data = wqw) +
    geom_boxplot( alpha = .5)+
    geom_jitter( alpha = .3)  +
    geom_smooth(aes(quality-2, 
                    volatile.acidity),
                method = "lm", 
                se = FALSE,size=1)
summary(aov(quality ~ volatile.acidity, data=wqw))
```


as volatile.acidity increasing , quality have a slight decreasing trend.

ANOVA show the influence is significant

#### chlorides and quanlity

```{r}
ggplot(aes(factor(quality), 
            chlorides), 
        data = wqw) +
    geom_boxplot( alpha = .5)+
    geom_jitter( alpha = .3)  +
    geom_smooth(aes(quality-2, 
                    chlorides),
                method = "lm", 
                se = FALSE,size=2)
summary(aov(quality ~ chlorides, data=wqw))
```
higher quality tender to have lower chlorides.

ANOVA show the chlorides's influence on quality is significant


#### total.sulfur.dioxide and quanlity


```{r}
ggplot(aes(factor(quality), 
            total.sulfur.dioxide), 
        data = wqw) +
    geom_boxplot( alpha = .5)+
    geom_jitter( alpha = .3)  +
    geom_smooth(aes(quality-2, 
                    total.sulfur.dioxide),
                method = "lm", 
                se = FALSE,size=2)
summary(aov(quality ~ total.sulfur.dioxide, data=wqw))
```

higher quality wine tender to have lower chlorides total.sulfur.dioxide.

ANOVA also show the influence is significant

## 4 Multivariate Plots Section

###  quality  Vs alcohol and density

```{r}
## make a plot of density and alcohol faceted by quality
ggplot(aes(x = alcohol, 
           y = density), 
       data = wqw) +
      geom_point(alpha = 0.1, size = 2) +
      geom_smooth(method = "lm", se = FALSE,size=1)+
      facet_wrap(~quality)

library(plyr)
ddply(wqw, "quality", summarise, corr=cor(alcohol, density))
```

from the plot we can see , higher quality(such as 7~9  ) wine more likely to appear on high acohol low density area .

higher qualiyt wines' correlation between alcohol and density appear to be more close to -1 , indicate that high quality wine must have less chemical ingredient that may influen density other than alcohol.  

### quality Vs chlorides and total.sulfur.dioxide

```{r}

ggplot(data = wqw,aes(x=chlorides,y=total.sulfur.dioxide))+ 
  geom_jitter(alpha =0.1,color="blue")+
  xlim(quantile(wqw$chlorides,0.01),
       quantile(wqw$chlorides,0.99))+
  ylim(quantile(wqw$total.sulfur.dioxide,0.01),
       quantile(wqw$total.sulfur.dioxide,0.99))+
   geom_smooth(method = "lm", se = FALSE,size=1)+
  facet_wrap(~quality,ncol = 3)

ddply(wqw, "quality", summarise, corr=cor(chlorides, total.sulfur.dioxide))

```


low quality wines' chlorides and total.sulfur.dioxide have a wide distribution;
higher quality wines' chlorides and total.sulfur.dioxide more focused on both low area.

correlation's differ betwwen different quality groups do not show something valuable.

## 5 Building the Linear Model

```{r Building the Linear Model}

## bulid a linear Model 
library(memisc)
m1 <- lm(I(quality ~alcohol), data = wqw)
m2 <- update(m1, ~ . + density)
m3 <- update(m2, ~ . + volatile.acidity)
m4 <- update(m3, ~ . + chlorides)
m5 <- update(m4, ~ . + total.sulfur.dioxide)
mtable(m1, m2, m3, m4, m5)
```


#### test our model

```{r}

### randam choose a 
a = subset(wqw,X == 2345)
q_pre <-
  round(predict(m5,newdata = a,interval="prediction", level = .95))[1,1]
q<-a$quality
q==q_pre
```

our model looks good!

## Final Plots and Summary


### plot 1 

Previous research find out that the distribution of wine in alcohol was multimodal. I want to take a close look at its distribution 

```{r}
ggplot(wqw,aes(alcohol))+
  geom_histogram(binwidth = 0.1,fill="blue")+
  scale_x_continuous(limits =c(quantile(wqw$alcohol,0.01),
                      quantile(wqw$alcohol,0.99)),
                      breaks = seq(8,14,0.5))+
  ggtitle("Alcohol Distribution")+ 
  xlab("Alcohol(mg/L)")+
  theme_minimal()

### cut alcohol 
wqw$alcohol_level=cut(wqw$alcohol,breaks = seq(8.5,13.5,0.5))
table(wqw$alcohol_level)

```


### Description 1
The distribution of alcolhol appears to be multimodal



from the summary , we can see there are 3 distribution peak at alcohol level about 9.5,10.5,12,5 mg/L , it seem that there may be 3 main kinds of  wine (high alcohol,low alcohol,mid alcohol)

### plot 2


```{r}
ggplot(aes(factor(quality), 
            total.sulfur.dioxide), 
        data = wqw) +
    geom_boxplot( alpha = .5)+
    geom_jitter( alpha = .3)  +
    geom_smooth(aes(quality-2, 
                    total.sulfur.dioxide),
                method = "lm", 
                se = FALSE,size=2)+
  ggtitle("Quality Vs total sulfur dioxide")+
  xlab("quality")+
  ylab("total sulfur dioxide(mg/L)")

ddply(wqw, "quality",summarise,
      mean=mean(total.sulfur.dioxide),
      median=median(total.sulfur.dioxide))
```

as we can see in the plot , there is  a tender that high quality wine have low content of sulfur dioxide.it can be justified by the mean and median of sulfur dioxide was lower in higher quality wine.




### plot 3
in the correlation matrix section ,we  find out that the largest correlation are correlation between density and alcohol

so I want to know how the relationship between density and alcohol influence wine quality



```{r}
library(plyr)
ggplot(aes(x = density, 
           y = alcohol),
           data = wqw) +
       ylim(8.5,quantile(wqw$alcohol,0.99))+
       xlim(0.987,quantile(wqw$density,0.99))+
       geom_point(aes(color = factor(quality)),
                  alpha = 0.05, size = 2) +
       geom_smooth(aes(color = factor(quality)),
                   method = "lm", se = FALSE,size=1)+
      ggtitle("density by alcohol")+
      xlab("density (g/ml)")+
      ylab("alcohol content (mg/L)")

  
ddply(wqw, "quality", summarise, 
      cor_alcohol_density=cor(alcohol, density))
```

### Description 3
 high quality wine more likely to appear in low alcohol/density area.
 
 higher quality wines' correlation between alcohol and density appear to be more  close to -1 , indicate that high quality wine must have less chemical ingredient  that may influen density other than alcohol.
 

 
 



## Reflection

1.  my vocabulary is very poor ,it take time to find proper word to discribe
2.  Quality level is int make it more like categorical variable , make it more diffcult to plot.
3. I am a total beginner in R .so it take me a lot of time to fix bugs 
3. future work : after I learn about logistic regression ,I can creat a better model.
